A comparative study of batch construction strategies \\ for recurrent neural networks in MXNet

### 2017-04-18 paper/paper.tex
experiments

  We extracted 16-dimensional MFCC vectors from beam-formed versions of the six sub-corpora in the training data corresponding to each channel
  and used them as features for the neural network models \cite{menne16:chime4System}.
  A CART tying was used to cluster the (roughly 120k possible) allophones into 1500 CART labels.
  We trained a three state HMM without skip transitions in the standard fashion and applied it in a forced alignment mode to the training data
  in order to obtain a frame-wise labeling of the data.

  \section{Experiments} \label{sec:experiments}
   We conducted experiments with different batch construction strategies. The results are reported in \ref{tab:chime:batch}, where the first three rows show results for trivial sequence ordering 
   methods. The last two rows then provide a direct comparison of the bucketing approach as 
   it is implemented in MXNet and the sinusoidal approach as proposed in this paper.
    %\begin{table}[tbp]
    %\centering
    %\caption{Comparison of runtime and memory requirement for different software packages. The %numbers were averaged over 100 training epochs. 
    %Note that sharp memory usage estimates for Torch and TensorFlow can not be obtained due to %their 5internal memory management.}
    %\label{tab:chime:compare}
    %\begin{tabular}{lrr}
    %  \hline
    %  Toolkit                 & Runtime [sec] & Memory [GB] \\
    %  \hline
    %  RETURNN   &  140     & 0.6     \\
    %  \hline
    %  Theano (in RETURNN)     &  370     & 0.8     \\
    %  TensorFlow              &  580     & 2.1$^*$     \\
    %  \hline
    %\end{tabular}
    %\end{table}

	\begin{table}[tbp]
		\centering
		\caption{Training time and recognition performance when training is done 
		on sub-sequence level. The first column shows the maximal sub-sequence length 
		after partitioning of the original sequence. The last line shows the results 
		if no partitioning into sub-sequence is performed. }
		\label{tab:chime:batch}
		\begin{tabular}{lrrr}
			\hline
			Chunk size              & Runtime [sec] & Memory [GB] & WER [\%] \\
			\hline
			10     					&				&			  &			 \\
			50  					&				&			  & 		 \\
			100 					& 				&			  &			 \\
			500						& 				&			  &			 \\
			$\max$					&				&			  &			 \\
			\hline
		\end{tabular}
	\end{table}

	\begin{table}[tbp]
		\centering
		\caption{An evaluation of different sequence batch construction methods on the CHiME-4 database. Training time per epoch, memory consumption are presented in the first two columns, while the last columns shows the word error rate of the corresponding system
		when decoded with the RASR speech recognition toolkit.}
		\label{tab:chime:batch}
		\begin{tabular}{lrrr}
			\hline
			Approach                & Runtime [sec] & Memory [GB] & WER [\%] \\
			\hline
			Static					&				&			  &			 \\
			Random 					&				&			  & 		 \\
			Sorted					& 				&			  &			 \\
			\hline
			Bucketing				&				&			  &			 \\
			Sinusoidal  			&				&			  &			 \\
			\hline
		\end{tabular}
	\end{table}

  As expected, sorting the entire training set by sequence length reduces the required time 
  per epoch to a minimum.

  \section{Conclusions}
    In this work we presented a novel strategy to construct sequence-level batches for recurrent 
    neural network training. Most system rely on a bucketing approach by clustering sequences of similar length into bins and to create batches from each bin individually. We showed that we can achieve a better runtime performance using a simpler batch design.

>>>>

  We extracted 16-dimensional MFCC vectors from beam-formed versions of the six sub-corpora in the training data corresponding to each channel
  and used them as features for the neural network models \cite{menne16:chime4System}.
  A CART tying was used to cluster the (roughly 120k possible) allophones into 1500 CART labels.
  We trained a three state HMM without skip transitions in the standard fashion and applied it in a forced alignment mode to the training data in order to obtain a frame-wise labeling of the data. A three layer 
  bi-directional LSTM followed a softmax layer was trained to minimize the cross-entropy of the softmax output and the CART label sequence at each frame. Optimization was done with the Adam optimizer and a constant learning rate of 0.01.

  After training, state posteriors were extracted from the softmax for each observation of the evaluation set. The posteriors were then decoded in the RWTH 
  speech recognition toolkit RASR \cite{rybach2011:rasr} using the hybrid HMM approach. A 4-gram LM 
  was used during decoding with a language model scale of 12.

  \section{Experiments} \label{sec:experiments}
   In order to provide some perspective on the required context on the CHiME-4 task, 
   we provide results for experiments on sub-utterance (chunk) in 
   Table \ref{tab:chime:chunk}. In the first columns we evaluate different sub-sequence lengths, while measuring the utterances that were processed per second, the memory required and the word error rate (WER) on the evaluation set of the CHiME-4 database. We can observe that 
   while large speed-ups can be obtained when training is done in this fashion, full-utterance context is
   required for optimal performance.

   For training on full sequences we conducted experiments with different batch construction strategies.
   The results are reported in Table \ref{tab:chime:batch}, where the first three rows show results for trivial sequence ordering 
   methods. The last two rows then provide a direct comparison of the bucketing approach as 
   it is implemented in MXNet and the alternated sorting approach as proposed in this paper.

	\begin{table}[tbp]
		\centering
		\caption{Training time and recognition performance when training is done 
		on sub-sequence level. The first column shows the maximal sub-sequence length 
		after partitioning of the original sequence. The last line shows the results 
		if no partitioning into sub-sequence is performed. }
		\label{tab:chime:chunk}
		\begin{tabular}{lrrr}
			\hline
			Chunk size              & Utt./sec & Memory [GB] & WER [\%] \\
			\hline
			10     					&				 &			   &			 \\
			50  					&				 &			   &    		 \\
			100 					& 				 &			   &			 \\
			500						& 				 &			   &			 \\
			$\max$					&				 &			   &			 \\
			\hline
		\end{tabular}
	\end{table}

	\begin{table}[tbp]
		\centering
		\caption{An evaluation of different sequence batch construction methods on the CHiME-4 database. Training time per epoch, memory consumption are presented in the first two columns, while the last columns shows the word error rate of the corresponding system
		when decoded with the RASR speech recognition toolkit.}
		\label{tab:chime:batch}
		\begin{tabular}{lrrr}
			\hline
			Approach                & Utt./sec & Memory [GB] & WER [\%]          \\
			\hline
			Random 					&	       & 			 & 6.8		 \\
			Sorted					& 		   &			 & 7.4		 \\
			\hline
			Bucketing				&		   &			 & 7.2		 \\
			Proposed (4 bins)	&		   &			 & 7.0		 \\
			Proposed (64 bins)	&		   &			 & 7.0		 \\
			Proposed (256 bins)	&		   &			 & 6.9		 \\
			\hline
		\end{tabular}
	\end{table}

  As expected, sorting the entire training set by sequence length reduces the required time 
  per epoch to a minimum, while the best overall performance is obtained when sequences are shuffled randomly.

  \section{Conclusions}
    In this work we presented a novel strategy to construct sequence-level batches for recurrent 
    neural network training. Most frameworks rely on a bucketing approach by clustering sequences of similar length into bins and to create batches from each bin individually. We showed that we can achieve a better runtime performance using a simpler batch design, by partitioning a shuffled sequence order and to sort the partitions in alternating order. The method was evaluated on the ChiME-4 noisy speech recognition task and 
    compared to the bucketing approach of MXNet.
